Report on the Second Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE2)

### 2015-05-27 WSSSPE2_report/WSSSPE2_review.tex
still working on 10

This group began by recognizing that they know what the code testing and code review problems are, so the purpose of the group is to think about concrete things that people in the group are willing to commit to, to improve sustainability of software.

Greg Wilson suggested that these concrete things should be small things so that the group members were likely to do them.

@@

The group asked if it could produce a 100-line example of how to test neutron transport (or other specific scientific examples) targeted at a sophomore? This would be a demonstration of ``how do I do this right''

Another question was ``How do I test so that I know that the code is not the source of my problems?''
A challenge is stripping things down to that level / that condensed
Test for infrastructure easier than models
cf. Lorena Barba’s “12 Steps to Navier Stokes” http://lorenabarba.com/blog/cfd-python-12-steps-to-navier-stokes/
And explanation of why we don’t teach testing in Software Carpentry any more http://software-carpentry.org/blog/2014/10/why-we-dont-teach-testing.html 
need to teach people how to test short bits of code, can’t wait until we have 1000s of lines of code
Distinction between 100 lines of input files vs. 100 lines of source code
Paper on multiphysics software verificationt01
"If we can't teach people how to test code until they're developing 1000s of lines of code, we can't be surprised when the don't write tests" - Greg Wilson
Don’t accept patches that don’t have tests
Setting up infrastructure to test well is difficult
Want a standard way that runs out of the box
Q: what about Jenkins?
A: sure, but you have to set it up, and that’s not easy for most scientists
People who’ve worked with a testing framework once (even if they have to be forced to do so) won’t work without it afterward
Q: but how to get them to do it the first time?
Two important pieces of the barrier:
Picking/implementing testing frameworks (technical barrier)
What are the actual tests that I write?
Ask students “how is what you’re doing different from typing random keys on the keyboard?” Turn their answer into the concept of tests.
How many scientists are aware of the fact that “testing the code” and “testing the science” are separate?
“Every scientist should go through a programming class”
That’s probably too big a goal for this group for the next year
Software Carpentry exists because university curricula are full; can’t add extra classes
Validation questions: how do you know that your method produces a result that is ‘close enough’
Tom Clune wants a piece of code that badly engineered/lacks unit tests so that he can add the unit tests and show people what it should look like
The collaboration must be with someone that either wrote the code or is intimately familiar with the topic.  The exercise will be to extract unit-tests through reverse-engineering {\b and} questioning the developer/expert
Code does not have to be “badly engineered”, but is either representative or is believed to be impenetrable from a unit-testing perspective.

>>>>

This group began by recognizing that they know what the code testing and code review problems are, so the purpose of the group was really to think about concrete things that people in the group would be willing to commit to, in order to improve the sustainability of software.  It was suggested that these concrete things should be small things so that the group members were likely to do them.

@@

The group asked if it could produce a 100-line example of how to test neutron transport (or other specific scientific examples) targeted at a sophomore? This would be a demonstration of ``How do I do this right?''

Another question was ``How do I test so that I know that the code is not the source of my problems?''
A challenge is stripping things down to an appropriate level.

Testing for infrastructure is easier than testing for models. \katznote{to delete? - don't see how it fits}

The group referred to Lorena Barba's ``12 Steps to Navier Stokes~\cite{12steps},'' an online CFD lesson repository created as a set of iPython notebooks.

Grid Wilson pointed to an explanation of why Software Carpentry doesn't teach testing any more~\cite{SCtesting}, which boils down to two issues.  First, that scientific computing doesn't (yet) have the cultural norms for error bars that experimental sciences have, and second, that there is a breathtaking diversity of scientific code; scientific research is highly specialized, which means that the tests scientists write are much less transferable or reusable than those found in other fields.

The group recognized that we need to teach people how to test short bits of code, we can't wait until we have thousands of lines of code. Greg commented, ``if we can't teach people how to test code until they're developing thousands of lines of code, we can't be surprised when they don't write tests.''

Distinction between 100 lines of input files vs. 100 lines of source code \katznote{to delete? - don't see how it fits}

in progress....

%Paper on multiphysics software verificationt01 - \url{http://onlinelibrary.wiley.com/doi/10.1002/spe.2220/abstract}
%
%Don't accept patches that don't have tests
%
%Setting up infrastructure to test well is difficult
%
%Want a standard way that runs out of the box
%Q: what about Jenkins?
%A: sure, but you have to set it up, and that's not easy for most scientists
%
%People who've worked with a testing framework once (even if they have to be forced to do so) won't work without it afterward
%Q: but how to get them to do it the first time?
%
%Two important pieces of the barrier:
%a. Picking/implementing testing frameworks (technical barrier)
%b. What are the actual tests that I write?
%Ask students ``how is what you're doing different from typing random keys on the keyboard?'' Turn their answer into the concept of tests.
%
%How many scientists are aware of the fact that ``testing the code'' and ``testing the science'' are separate?
%
%``Every scientist should go through a programming class''
%a. That's probably too big a goal for this group for the next year
%b. Software Carpentry exists because university curricula are full; can't add extra classes
%
%
%Validation questions: how do you know that your method produces a result that is ‘close enough’
%
%Tom Clune wants a piece of code that badly engineered/lacks unit tests so that he can add the unit tests and show people what it should look like
%a. The collaboration must be with someone that either wrote the code or is intimately familiar with the topic.  The exercise will be to extract unit-tests through reverse-engineering {\b and} questioning the developer/expert
%b. Code does not have to be ``badly engineered'', but is either representative or is believed to be impenetrable from a unit-testing perspective.
